Calculating Module Enrichment and Visualizing Data on Large-scale Molecular Maps with the R packages \pkg{ACSNMineR} and \pkg{RNaviCell}

### 2015-11-25 jss/article.tex
jss update

In this article, we describe two software packages related to ACSN analysis and
data visualization for the popular R statistical enrironment \citep{mainRref,
vance2009data}. The package \pkg{ACSNMineR} is designed for the calculation of
gene enrichment and depletion in ACSN maps, while \pkg{RNaviCell} is dedicated
to data visualization on ACSN maps. Both packages are available on the
Comprehensive R Archive Network
(\url{https://cran.r-project.org/web/packages/ACSNMineR/} and
\url{https://cran.r-project.org/web/packages/RNaviCell/}), and on the GitHub
repository (\url{https://github.com/sysbio-curie/ACSNMineR} and
\url{https://github.com/sysbio-curie/RNaviCell}).

>>>>

In this article, we describe two software packages related to ACSN analysis and
data visualization for the popular R statistical enrironment \citep{mainRref,
vance2009data}. The package \pkg{ACSNMineR} is designed for the calculation of
gene enrichment and depletion in ACSN maps, while \pkg{RNaviCell} is dedicated
to data visualization on ACSN maps. Both packages are available on the
Comprehensive R Archive Network
(\url{https://cran.r-project.org/web/packages/ACSNMineR/} and
\url{https://cran.r-project.org/web/packages/RNaviCell/}), and on the GitHub
repository (\url{https://github.com/sysbio-curie/ACSNMineR} and
\url{https://github.com/sysbio-curie/RNaviCell}). For the remainder of this
article, we describe the organization of each package and illustrate their
capacities with several concrete examples demonstrating their capabilities.

\section[Package organization]{Packages organization}

\subsection{ACSNMineR}
Currently, ACSN maps covers signaling pathways involved in DNA repair, cell cycle,
cell survival, cell death, epithelial-to-mesenchymal transition (EMT) and cell
motility. The information included in ACSN is organized in three hierarchical
levels: a global map, which is composed of five individual maps, themselves
composed of many functional modules covering individual pathways. Each map in
ACSN covers hundreds of molecular players, biochemical reactions and causal
relationships between the molecular players and cellular phenotypes. ACSN
represents a large-scale biochemical reaction network of 4,826 reactions
involving 2,371 proteins, and is continuously updated and expanded. 
We have included the three hierarchical levels in the \pkg{ACSNMineR} package,
in order to be able to calculate enrichments at all three levels.

\subsection{RNaviCell}
